---
title: "Attrition Case Study"
author: "Jae Chung"
date: "2023-04-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Loaded in every library needed for this analysis
```{r}
library(ggplot2)
library(stringi)
library(readxl)
library(XML)
library(rvest)
library(naniar)
library(e1071)
library(caret)
library(class)
library(tidyverse)
library(GGally)
library(stringr)
library(dplyr)
library(tidyr)
library(knitr)
library(rmarkdown)
library(psych)
library(Metrics)
library(plotly)

```

#Imported CaseStudy2-data.csv file
```{r}
#Reading in employee data
employee = read.csv(file.choose(), header = TRUE)
employee

#Verifying employee data
head(employee)
```

#Cleaned data for potential missing values
```{r}
#Removing NAs in dataset
is.na(employee)
employeeClean = na.omit(employee)

#Summarizing data
str(employeeClean)
summary(employeeClean)
```

#Changed OverTime to binary category
```{r}
#Converting the Overtime fields to a binary category for scalability
employeeClean$OverTimeBinary = ifelse(employeeClean$OverTime =="Yes", 1,0) 
```

#Scaled columns for comparison of multiple variables
```{r}
#Scaling columns
employeeClean$MonthlyIncomeScale = scale(employeeClean$MonthlyIncome)
employeeClean$AgeScale = scale(employeeClean$Age)
employeeClean$EnvironmentSatisfactionScale = scale(employeeClean$EnvironmentSatisfaction)
employeeClean$OverTimeScale = scale(employeeClean$OverTimeBinary)
employeeClean$YearsAtCompanyScale = scale(employeeClean$YearsAtCompany)

#Verifying scale
employeeClean
head(employeeClean)

```

#Plotted correlation between Attrition and other columns
```{r}
#Attrition Count
ggplotly(employeeClean %>% ggplot(aes(x = Attrition, fill = Attrition)) + geom_bar() + labs(title = "Attrition Count", x = "Attrition"))

#Attrition Vs. Age
ggplotly(employeeClean %>% ggplot(aes(x = Age, fill = Attrition)) + geom_histogram() + ggtitle("Attrition Vs. Age"))

#Attrition vs. Business Travel
ggplotly(employeeClean %>% ggplot(aes(x = BusinessTravel, fill = Attrition)) + geom_bar(position = "fill") + ggtitle("Attrition Vs. Business Travel"))

#Categorizing Daily Rates into groups
min(employeeClean$DailyRate)
max(employeeClean$DailyRate)
employeeClean$DailyRateGroups <- as.factor(ifelse(employeeClean$DailyRate<501, "0-500", ifelse(employeeClean$DailyRate<1001, "501-1000", ifelse(employeeClean$DailyRate<1501, "1001-1500"))))

#Attrition vs. Daily Rate
ggplotly(employeeClean %>% ggplot(aes(x = DailyRateGroups, fill = Attrition)) + geom_bar() + ggtitle("Attrition Vs. Daily Rate"))

#Attrition vs. Department
ggplotly(employeeClean %>% ggplot(aes(x = Department, fill = Attrition)) + geom_bar() + ggtitle("Attrition Vs. Department"))

#Categorizing Distance from Home into groups
min(employeeClean$DistanceFromHome)
max(employeeClean$DistanceFromHome)
employeeClean$DistanceFromHomeGroups <- as.factor(ifelse(employeeClean$DistanceFromHome<11, "0-10", ifelse(employeeClean$DistanceFromHome<21, "11-20", ifelse(employeeClean$DistanceFromHome<31, "21-30"))))

#Attrition vs. Distance From Home
ggplotly(employeeClean %>% ggplot(aes(x = DistanceFromHomeGroups, fill = Attrition)) + geom_bar() + ggtitle("Attrition Vs. Distance From Home"))

#Converting Education into Factors
employeeClean$EducationFactor <- as.factor(employeeClean$Education)

#Attrition vs. Education
ggplotly(employeeClean %>% ggplot(aes(x = EducationFactor, fill = Attrition)) + geom_bar() + ggtitle("Attrition Vs. Education"))

#Converting Education Field into Factors
employeeClean$EducationFieldFactor <- as.factor(employeeClean$EducationField)

#Attrition vs. Education Field
ggplotly(employeeClean %>% ggplot(aes(x = EducationFieldFactor, fill = Attrition)) + geom_bar() + ggtitle("Attrition Vs. Education Field"))

#Converting Environment Satisfaction into Factors
employeeClean$EnvironmentSatisfactionFactor <- as.factor(employeeClean$EnvironmentSatisfaction)

#Attrition vs. Environment Satisfaction
ggplotly(employeeClean %>% ggplot(aes(x = EnvironmentSatisfactionFactor, fill = Attrition)) + geom_bar() + ggtitle("Attrition Vs. Environment Satisfaction"))

#Attrition vs. Gender
ggplotly(employeeClean %>% ggplot(aes(x = Gender, fill = Attrition)) + geom_bar() + ggtitle("Attrition Vs. Gender"))

#Categorizing Hourly Rates into groups
min(employeeClean$HourlyRate)
max(employeeClean$HourlyRate)
employeeClean$HourlyRateGroups <- as.factor(ifelse(employeeClean$HourlyRate<56, "0-55", ifelse(employeeClean$HourlyRate<81, "56-80", ifelse(employeeClean$HourlyRate<106, "81-105"))))

#Attrition vs. Hourly Rates
ggplotly(employeeClean %>% ggplot(aes(x = HourlyRateGroups, fill = Attrition)) + geom_bar() + ggtitle("Attrition Vs. Hourly Rates"))

#Attrition vs. Job Involvement
ggplotly(employeeClean %>% ggplot(aes(x = JobInvolvement, fill = Attrition)) + geom_bar() + ggtitle("Attrition Vs. Job Involvement"))

#Attrition vs. Job Level
ggplotly(employeeClean %>% ggplot(aes(x = JobLevel, fill = Attrition)) + geom_bar(position = "fill") + ggtitle("Attrition Vs. Job Level"))

#Attrition vs. Job Role
ggplotly(employeeClean %>% ggplot(aes(x = JobRole, fill = Attrition)) + geom_bar(position = "fill") + ggtitle("Attrition Vs. Job Role"))

#Attrition Vs. Job Satisfaction
ggplotly(employeeClean %>% ggplot(aes(x = JobSatisfaction, fill = Attrition)) + geom_bar() + ggtitle("Attrition Vs. Job Satisfaction"))

#Attrition Vs. Marital Status
ggplotly(employeeClean %>% ggplot(aes(x = MaritalStatus, fill = Attrition))+ geom_bar() + ggtitle("Attrition Vs. Marital Status"))

#Attrition Vs. Monthly Income
ggplotly(employeeClean %>% ggplot(aes(x = MonthlyIncome, fill = Attrition)) + geom_histogram() + ggtitle("Attrition Vs. Monthly Income"))

#Attrition Vs. Monthly Rate
ggplotly(employeeClean %>% ggplot(aes(x = MonthlyRate, fill = Attrition)) + geom_histogram() + ggtitle("Attrition Vs. Monthly Rate"))

#Attrition Vs. Number of Companies Worked
ggplotly(employeeClean %>% ggplot(aes(x = NumCompaniesWorked, fill = Attrition)) + geom_histogram() + ggtitle("Attrition Vs. Number of Companies Worked"))

#Attrition Vs. Overtime
ggplotly(employeeClean %>% ggplot(aes(x = OverTime, fill = Attrition)) + geom_bar(position = "fill") + ggtitle("Attrition Vs. Overtime"))

#Attrition Vs. Percent of Salary Hike
ggplotly(employeeClean %>% ggplot(aes(x = PercentSalaryHike, fill = Attrition)) + geom_bar() + ggtitle("Attrition Vs. Percent of Salary Hike"))

#Converting Performance Rating into Factors
employeeClean$PerformanceRatingFactor <- as.factor(employeeClean$PerformanceRating)

#Attrition Vs. Performance Rating
ggplotly(employeeClean %>% ggplot(aes(x = PerformanceRatingFactor, fill = Attrition)) + geom_bar() + ggtitle("Attrition Vs. Performance Rating"))

#Attrition Vs. Relationship Satisfaction
ggplotly(employeeClean %>% ggplot(aes(x = RelationshipSatisfaction, fill = Attrition)) + geom_bar(position = "fill") + ggtitle("Attrition Vs. Relationship Satisfaction"))

#Attrition Vs. Stock Option Level
ggplotly(employeeClean %>% ggplot(aes(x = StockOptionLevel, fill = Attrition)) + geom_bar(position = "fill") + ggtitle("Attrition Vs. Stock Option Level"))

#Attrition Vs. Total Working Years
ggplotly(employeeClean %>% ggplot(aes(x = TotalWorkingYears, fill = Attrition)) + geom_bar(position = "fill") + ggtitle("Attrition Vs. Total Working Years"))

#Attrition Vs. Training Times Last Year
ggplotly(employeeClean %>% ggplot(aes(x = TrainingTimesLastYear, fill = Attrition)) + geom_histogram() + ggtitle("Attrition Vs. Training Times Last Year"))

#Attrition Vs. Work Life Balance
ggplotly(employeeClean %>% ggplot(aes(x = WorkLifeBalance, fill = Attrition)) + geom_histogram() + ggtitle("Attrition Vs. Work Life Balance"))

#Attrition Vs. Years at the Company
ggplotly(employeeClean %>% ggplot(aes(x = YearsAtCompany, fill = Attrition)) + geom_histogram() + ggtitle("Attrition Vs. Years At the Company"))

#Attrition Vs. Years in Current Role
ggplotly(employeeClean %>% ggplot(aes(x = YearsInCurrentRole, fill = Attrition)) + geom_histogram() + ggtitle("Attrition Vs. Years in Current Role"))

#Attrition Vs. Years Since Last Promotion
ggplotly(employeeClean %>% ggplot(aes(x = YearsSinceLastPromotion, fill = Attrition)) + geom_histogram() + ggtitle("Attrition Vs. Years Since Last Promotion"))

#Attrition Vs. Years with Current Manager
ggplotly(employeeClean %>% ggplot(aes(x = YearsWithCurrManager, fill = Attrition)) + geom_histogram() + ggtitle("Attrition Vs. Years with Current Manager"))

#Job Satisfaction Vs. Job Involvement
ggplotly(employeeClean %>% ggplot(aes(x = JobSatisfaction, y = JobInvolvement, color = Attrition)) + geom_point(position = "jitter") + ggtitle("Job Satisfaction Vs. Job Involvement"))

#Job Role Vs. Job Satisfaction
ggplotly(employeeClean %>% ggplot(aes(x = JobRole, y = JobSatisfaction, color = Attrition)) + geom_point(position = "jitter") + ggtitle("Job Role Vs. Job Satisfaction"))

#Job Role Vs. Work Life Balance
ggplotly(employeeClean %>% ggplot(aes(x = JobRole, y = WorkLifeBalance, color = Attrition)) + geom_point(position = "jitter") + ggtitle("Job Role Vs. Work Life Balance"))

```

#Ran test to find K 
```{r}
# Loop for many k and the average of many training / test partition
set.seed(4)

iterations = 500
numks = 30
masterAcc = matrix(nrow = iterations, ncol = numks)

for(j in 1:iterations)
{
  accs = data.frame(accuracy = numeric(30), k = numeric(30))
  trainIndices = sample(1:dim(employeeClean)[1], round(0.70* dim(employeeClean)[1]))
  train = employeeClean[trainIndices,]
  test = employeeClean[-trainIndices,]
  for(i in 1:numks)
  {
    classifications = knn(train[,c(37,38,40)], test[,c(37,38,40)], train$Attrition, prob = TRUE, k = i)
    table(classifications, test$Attrition)
    cm = confusionMatrix(table(classifications, test$Attrition))
    masterAcc[j,i] = cm$overall[1]
  }
}
MeanAcc = colMeans(masterAcc)
plot(seq(1,numks,1),MeanAcc, type = "l")

```

#KNN Model
```{r}
#KNN
classifications = knn(train[,c(37,38,40)], test[,c(37,38,40)], train$Attrition, prob = TRUE, k = 15) 
classifications

table(classifications, test$Attrition)
cm = confusionMatrix(table(classifications, test$Attrition))
cm
# Accuracy : 0.8621, Sensitivity : 0.9381, Specificity : 0.3714  
```

#Naive Bayes
```{r}
#Selecting columns that contribute to Attrition
model = employeeClean %>% select(c("MonthlyIncome","JobRole","Age","OverTime","EnvironmentSatisfaction","DailyRate","JobLevel","TotalWorkingYears","JobInvolvement","Attrition"))

trainIndices = sample(1:dim(model)[1],round(.70 * dim(model)[1]))
train = model[trainIndices,]
test = model[-trainIndices,]

set.seed(7)
classifier1 = naiveBayes(model[,c(1,4,7)],model$Attrition)
pred = predict(classifier1,newdata=test)
CM = confusionMatrix(table(test$Attrition,pred))

CM

```

#Found useful correlations  
```{r}
cor(employeeClean$MonthlyIncome ,employeeClean[,c(2,5,7,8,12,14,15,16,18,20,21,25,26,27,29,30, 31,32,33,34,35,36,37)])

```

#Linear Regression
```{r}
#Changing Attrition to numeric
employeeClean$AttritionInt = if_else(employeeClean$Attrition =="Yes", 1,0)

employeeClean$OverTimeFix = ifelse(employeeClean$OverTime=="Yes",1,0)

fit = lm(AttritionInt~MonthlyIncome + OverTimeFix + JobLevel, employeeClean)

summary(fit)
confint(fit)
plot(fit)
AIC(fit)

```

#Imported No Salary Data
```{r}
employeeSalary = read.csv(file.choose(), header = TRUE)
```

#Cleaned No Salary Data
```{r}
employeeSalaryClean = na.omit(employeeSalary)

head(employeeSalaryClean)
employeeSalaryClean

```
```{r}
cor(employeeClean$MonthlyIncome ,employeeClean[,c(2,5,7,8,12,14,15,16,18, 20,21,25,26,27,29,30,31,32,33,34,35,36,37)])

```

#RMSE
```{r}
set.seed(7)

LMSalaryMod = lm(MonthlyIncome ~ JobLevel + Age + TotalWorkingYears + YearsAtCompany + WorkLifeBalance + YearsWithCurrManager + DistanceFromHome + DailyRate + StockOptionLevel + PerformanceRating+HourlyRate + OverTime + YearsSinceLastPromotion + RelationshipSatisfaction + YearsInCurrentRole + PercentSalaryHike + TrainingTimesLastYear + MonthlyRate + JobSatisfaction + JobInvolvement, data = employeeClean)

summary(LMSalaryMod)

rss = c(crossprod(LMSalaryMod$residuals))
mse = rss / length(LMSalaryMod$residuals)
rmse = sqrt(mse)
sig2 = rss / LMSalaryMod$df.residual

rmse

```

#Predicted Salary to include in file
```{r}
SalaryPredict = predict(LMSalaryMod, newdata = employeeSalaryClean)
SalaryPredict
```

#Wrote Predicted Salary to file 
```{r}
employeeSalaryClean$PredictedSalary = unlist(SalaryPredict)
PredictionSalaryDF = data.frame(c(employeeSalaryClean$ï..ID), c(employeeSalaryClean$PredictedSalary))

SalaryPrediction = unlist(SalaryPredict)

employeeSalary$SalaryPrediction = SalaryPrediction
employeeSalary$SalaryPrediction
head(employeeSalary)

write.csv(PredictionSalaryDF, "SalaryPrediction.csv", row.names = TRUE)

```

#Predicted Attrition and wrote to file
```{r}
employeeNoAttri = read.csv(file.choose(), header = TRUE) 

classifier1 = naiveBayes(model[,-c(1,5,9,10,11,14,21,23,24,28, 37:45)],model$Attrition)

attrition_pred = predict(classifier1, employeeClean)
employeeClean$PredictAttrition = unlist(attrition_pred)

PredictionsAttritionDF = data.frame(c(employeeClean$ID), c(employeeClean$PredictAttrition))

write.csv(PredictionsAttritionDF, "AttritionPrediction.csv", row.names = TRUE)

```
